The E-C Algorithm: conceptual bootstrapping via grammar-based program compression.

### 2013-01-21 drafts/ijcai2013/ijcai13.tex
commandline arugments for symbolic regressioin experiment.

\begin{figure}
  \centering
  \begin{subfigure}[Before]{0.5\linewidth}
    \begin{tikzpicture}[scale=.7,
        transform shape, level/.style={sibling distance=10mm/#1, font=\scriptsize}
        ]
      \node [font=\scriptsize] (top) {$I$}
      child {node [above=.3cm] (a) {$(I \rightarrow I) \rightarrow I \rightarrow I$}
        child {node [left=.5cm, below=1cm, text width=3cm] 
          {$S : \\(t_2 \rightarrow t_1 \rightarrow t_0)  \rightarrow (t_2 \rightarrow t_1) 
            \rightarrow \\ t_2 \rightarrow t_0$}}
        child {node [right=0cm, text width=2cm] {$*: Int \rightarrow Int \rightarrow Int$}}
      }
      child {node [below right=.3cm and 1cm] (b) {$I \rightarrow I$}
        child {node {$I: t_0 \rightarrow t_0 \rightarrow t_0$}}
      };
    \end{tikzpicture}
    \caption{The typed combinator $S * I$ ($f(x)=x^2$) represented as binary tree.}
    \label{fig:clbintree}
  \end{subfigure}
  \begin{subfigure}[After]{0.4\linewidth}
    \begin{tikzpicture}[scale=.7, transform shape, 
        level/.style={sibling distance=10mm/#1},
        every edge/.style={draw, thin}
      ]
      \node [circle,draw] (a) {}
      child {node [left=0cm] (b) {$f_1$}}
      child {node [left=.2cm] (c) {\dots} edge from parent[draw=none] }
      child {node [left=.4cm] (e) {$f_n$}}
      child {node [circle,draw, left=1cm] (d) {} edge from parent[very thick]
        child {node [circle, draw, left=.25cm] (l) {} edge from parent[draw=none]
          child {node [left=.5cm] (b) {$f_1$}} 
          child {node [left=.3cm] (c) {\dots} edge from parent[draw=none] }
          child {node [left=.3cm] (e) {$f_n$} edge from parent[very thick]}
          child {node [circle, draw, left=.25cm] (and2) {}
            child {node [circle, draw, left=.25cm](l2) {} edge from parent[draw=none] 
              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
            }
            child {node [circle, draw] (r2)  {} edge from parent[draw=none] 
              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
            }
          }
        }
        child {node [circle, draw, right=.25cm] (r) {} edge from parent[draw=none]
          child {node [circle, draw, right=.25cm] (and3) {}
            child {node [circle, draw ](l3) {} edge from parent[draw=none] 
              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
            }
            child {node [circle, draw, right=.2cm ] (r3)  {} edge from parent[draw=none] 
              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
            }
          }
          child {node [right=.3cm] (b) {$f_1$} edge from parent[very thick]} 
          child {node [right=.4cm] (c) {\dots} edge from parent[draw=none] }
          child {node [right=.6cm] (e) {$f_n$}}
        }
      };

@@

      \path (and3)  edge node[inner sep=0mm,pos=.4] (t1) {} node {} (l3);
      \path (and3)  edge node[inner sep=0mm,pos=.4] (k1) {} node {} (r3);
      \path (t1) edge[bend right] (k1) ;
    \end{tikzpicture}
    \caption{The space of all typed combinators represented as an
      infinitely deep And-Or tree. Any specific combinator as a
      sub-binary tree of this And-Or tree whose leaves consist of only
      the $f_i$ nodes (i.e., the primitive combinators of the
      grammar).}
    \label{fig:andor}
    \end{subfigure}
  \caption{A typed binary tree representation of the combinator $S\;*\;I\;$.}
\end{figure}

@@

In our algorithm, we traverse a set of combinator binary trees in
depth first order. We count the occurences of every subtree. Every
time we encounter a subtree for the second time, we designate it as a
primitive in our new grammar. Furthermore, during our traversal, we do
not descend into trees that have been added already to the new
grammar. Thus, we only count an instance of tree if that instance is
not a subtree of a primitive element of the new grammar.

We want to count the number of
times every subtree is encountered when it is not a subtree of a tree
already encountered in our traversal. we simply count the numer of
times each unique subtree appears as the child of a subtree \emph{not
  yet encountered} in the solution set. That is, we traverse the
solution set in depth first order, comparing each tree we encounter to
all the trees we previously encountered; but if we encounter a tree we
have already seen, we do not descend into it. In this traversal, the
set of trees we encounter more than once is the set of terminal nodes
in our new grammar. To estimate the probabilities associated with each
node production, we again traverse the solution set, but these time
for each node where a terminal production takes place we count which
other terminal productions could have taken place. We estimate the
terminal production weight to be the ratio of the number of times that
node was used in the first traversal divided by the number of times it
was used in the second traversal (plus 1, to account for the event of
an $AND$ node expansion). Likewise, we re-estimate the probability of
an $AND$ node expansion by diving the number of such expansions in the
solution set by the number of possible expansions in the solution set.

>>>>

\begin{figure}
  \centering
  \begin{subfigure}[Before]{0.5\linewidth}
    \begin{tikzpicture}[scale=.9,
        transform shape, level/.style={sibling distance=10mm/#1, font=\scriptsize}
        ]
      \node [font=\scriptsize] (top) {$I$}
      child {node [above=.3cm] (a) {$(I \rightarrow I) \rightarrow I \rightarrow I$}
        child {node [left=.5cm, below=1cm, text width=3cm] 
          {$S : \\(t_2 \rightarrow t_1 \rightarrow t_0)  \rightarrow (t_2 \rightarrow t_1) 
            \rightarrow \\ t_2 \rightarrow t_0$}}
        child {node [right=0cm, text width=2cm] {$*: Int \rightarrow Int \rightarrow Int$}}
      }
      child {node [below right=.3cm and 1cm] (b) {$I \rightarrow I$}
        child {node {$I: t_0 \rightarrow t_0 \rightarrow t_0$}}
      };
    \end{tikzpicture}
    \label{fig:clbintree}
  \end{subfigure}
  \begin{subfigure}[After]{0.4\linewidth}
    \begin{tikzpicture}[scale=.7, transform shape, 
        level/.style={sibling distance=10mm/#1},
        every edge/.style={draw, thin}
      ]
      \node [circle,draw] (a) {}
      child {node [left=0cm] (b) {$f_1$}}
      child {node [left=.2cm] (c) {\dots} edge from parent[draw=none] }
      child {node [left=.4cm] (e) {$f_n$}}
      child {node [circle,draw, left=1cm] (d) {} edge from parent[very thick]
        child {node [circle, draw, left=.25cm] (l) {} edge from parent[draw=none]
          child {node [left=.5cm] (b) {$f_1$}} 
          child {node [left=.3cm] (c) {\dots} edge from parent[draw=none] }
          child {node [left=.3cm] (e) {$f_n$} edge from parent[very thick]}
          child {node [circle, draw, left=.25cm] (and2) {}
            child {node [circle, draw, left=.25cm](l2) {} edge from parent[draw=none] 
              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
            }
            child {node [circle, draw] (r2)  {} edge from parent[draw=none] 
              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
            }
          }
        }
        child {node [circle, draw, right=.25cm] (r) {} edge from parent[draw=none]
          child {node [circle, draw, right=.25cm] (and3) {}
            child {node [circle, draw ](l3) {} edge from parent[draw=none] 
              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
            }
            child {node [circle, draw, right=.2cm ] (r3)  {} edge from parent[draw=none] 
              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
            }
          }
          child {node [right=.3cm] (b) {$f_1$} edge from parent[very thick]} 
          child {node [right=.4cm] (c) {\dots} edge from parent[draw=none] }
          child {node [right=.6cm] (e) {$f_n$}}
        }
      };

@@

      \path (and3)  edge node[inner sep=0mm,pos=.4] (t1) {} node {} (l3);
      \path (and3)  edge node[inner sep=0mm,pos=.4] (k1) {} node {} (r3);
      \path (t1) edge[bend right] (k1) ;
    \end{tikzpicture}
    \label{fig:andor}
    \end{subfigure}
  \caption{(a) The typed combinator $S * I$ ($f(x)=x^2$) represented as
    binary tree. (b) The space of all typed combinators represented as an
    infinitely deep AND/OR tree. Any specific combinator as a
    sub-binary tree of this And-Or tree whose leaves consist of only
    the $f_i$ nodes (i.e., the primitive combinators of the grammar).}

\end{figure}

@@

In our algorithm, we traverse a set of combinator binary trees in
depth first order. We count the occurences of every subtree. Every
time we encounter a subtree for the \emph{second} time, we designate
it as a primitive in our new grammar. This ensures that every
primitive is used more than once. During our traversal, we do not
descend into trees that have been added already to the new
grammar. Thus, we only count an instance of tree if that instance is
not a subtree of a primitive element of the new grammar.

At the end of this traversal of the solution programs, we have a parse
of the solution set with a new set of primitives and we know how many
times each primitive occured in the solutions et. To estimate the
probabilities associated with each node production, we again traverse
the solution set, but this time for each node we count which other
primitive elements from the new grammar could have been used. We
estimate the terminal production weight to be the ratio of the number
of times that node was used in the first traversal divided by the
number of times it was used in the second traversal.

Note that this probability estimation procedure was chosen for its
simplicity, and it is not the maximum likelihood estimator for the
multinomial distribution according to which we enumerate the
data. Because the probability we assign to the use of a primitive
combinator depends on the type of the node, the ML estimator of the
probability of that combinator is not a simple ratio of counts. It
would be interesting to determine whether a more accurate estimator
can improve this algorithm's performance.
